Physics-informed Implicit Neural Representation for Potential Field Geophysics

### 2023-07-30 paper3.tex
Add physics sections and format

\documentclass{article}
\input{preamble}
\begin{document}

\author{\and{Luke Thomas Smith} \and{Tom Horrocks} \and {Eun-Jung Holden} \and{Daniel Wedge} \and{Naveed Akhtar}}
\title{Implicit Neural Representation Case Studies for Potential Field Geophysics}
\date{\today}
\maketitle{}

@@

\subsection{Implicit Neural Representation}
\label{sec:inr}
Realising their power in implicit representation for natural images, many works quicky investigated 
\begin{enumerate}
    \item MLPs
    \item Periodic activations
    \item Storage weights not cells
\end{enumerate}

\subsection{Geophysical Sampling}
\label{sec:geo_sampling}
Here, I discuss sampling in geophysics

\section{Method}
\subsection{SIREN}
\begin{enumerate}
    \item SIREN (or WIRE, if it improves)
    \item Regularisation loss
    \item Parameter opt
\end{enumerate}

@@

\section{Discussion}
\begin{enumerate}
    \item Limitations: Capacity of Network
    \item RAM usage (if I can't do batched C-MLP training)
    \item Accuracy of learnt x,y and z gradients
\end{enumerate}

>>>>

\documentclass{article}
\input{preamble}
\usepackage{amssymb}
\begin{document}

\author{\and{Luke Thomas Smith} \and{Tom Horrocks} \and {Eun-Jung Holden} \and{Daniel Wedge} \and{Naveed Akhtar}}
\title{Physics-informed Implicit Neural Representation for Potential Field Geophysics}
\date{\today}
\maketitle{}

@@

\subsection{Implicit Neural Representation}
\label{sec:inr}
Seminal work by \parencite{mildenhallNeRFRepresentingScenes2020} demonstrated the capacity for learning continuous functions from a set of coordinates.
However, it was not until they were combined with periodic activation functions \parencite{sitzmann2019siren} that the implicit representations could adequately recreate high frequency signal.
Following investigation by \parencite{ramasinghePeriodicityUnifyingFramework2022} shows that it is the magnitude of the first and second derivatives that enhance high-frequency capacity, rather than periodicity.

Realising their power in implicit representation for natural images, many works quicky investigated
\begin{enumerate}
    \item MLPs
    \item Periodic activations
    \item Storage weights not cells
\end{enumerate}

\subsection{Automatic Differentiation}
Automatic differentiation (AD) refers to one of several methods for computing derivatives, specifically to the accumulation of derivatives during function evaluation \parencite{baydin2018automatic}.
This method underpins modern machine learning frameworks such as Pytorch \parencite{paszkePyTorchImperativeStyle2019}.
In these frameworks, when evaluating a set of inputs against the models weight parameters, the derivatives with respect to the parameters are used to perform backpropogation when evaluating an objective function (loss).
These derivative inform weight updates in the direction of the negative gradient, toward a minima in the objective function (for example, L2 norm between predictions and targets).
However, automatic differentiation can also evaluate the derivatives with respect to the inputs.
The gradient provides partial derivatives $\frac{\partial f}{\partial x}$, $\frac{\partial f}{\partial y}$, $\frac{\partial f}{\partial z}$.

\subsection{Physics-Informed Neural Networks}
\parencite{raissiPhysicsinformedNeuralNetworks2019} introduced Physics-informed neural networks (PINNs).
When dealing with samples of phenomena which obey natural laws, there also exists \emph{a priori} constraints on those samples.
These constrains can be used to regularise neural networks, vastly reducing the search space for physically possible solutions.

\parencite{liImplicitStochasticGradient2023} replace the ubiquitous stochastic gradient descent optimiser \emph{Adam} with an implicit stochastic gradient descent (ISGD) method:
$$\theta_{n+1} = \theta_{n} - \alpha \cdot \nabla L (\theta_{n+1})$$

\parencite{sethiHardEnforcementPhysicsinformed2023} demonstrate that this regularisation does not have to be implemented as a loss function, but instead can be a \emph{hard} constraint embedded within the network.
This hard constraint takes the form of a function, which satisfies the \emph{a priori} laws, applied to the output of the network prior PDE and regularisation.

Potential fields in an airborne exploration context are constrained by a number of physical laws.
From \parencite{blakelyPotentialTheoryGravity1996}, potential fields satisfy \emph{Laplace's equation},
$$\nabla^2 \phi = 0$$
This indicates the function lacks curvature, in three dimensions $\phi(x,y,z)$:
$$\frac{\partial^2\phi}{\partial x^2} + \frac{\partial^2\phi}{\partial y^2} + \frac{\partial^2\phi}{\partial z^2} = 0$$

\subsection{Geophysical Sampling and Continuation}
\label{sec:geo_sampling}
Potential fields in geophysics are vector fields, but are typically sampled as a scalar in one direction.
A scalar field can be parameterised with a single dimension at each spatial coordinate, while a vector field requires three orthogonal components at each coordinate \parencite{blakelyPotentialTheoryGravity1996}.

$F = \nabla \phi$ for the gravity potential, $F = - \nabla \phi$ for the Magnetic potential.

$\phi(x,y,z) = constant$ is an equipotential surface. If Laplaces equation: $\nabla^2 \phi = 0$ is satisfied, it is harmonic, and has continuous single valued first derivatives, and has second derivatives.

\section{Method}
\subsection{Notation}
$\mathbb{R}^n$ Real n-dimensional vectors, named using lower-case letters, e.g. \boldmath{x}.

$\mathbb{R}^{m\times{}n}$ $m*n$ dimensional matrices, named using upper-case letters, e.g. \boldmath{A}.

\subsection{Differentiation}
A Numerical Differentiation approach to gradient calculation here would be to evaluate $\frac{\partial{U}}{\partial{(x, y, z)}}$ at a set of randomly sampled coordinates $(x, y, z)_1 ... (x,y,z)_n$.
However, using Automatic Differentiation within Pytorch, the continuous implicit represenation can be differentiated to provide gradients without being constrained by numerical precision or quantisation.
These are also unstable, recognised in the geophysical community by divergence and noise in predictions depending on numerical methods.

\subsection{SIREN}
\begin{enumerate}
    \item SIREN (or WIRE, if it improves)
    \item Regularisation loss
    \item Parameter opt
\end{enumerate}

@@

\section{Discussion}
\begin{enumerate}
    \item Limitations: Capacity of Network
    \item RAM usage (if I can't do batched C-MLP training)
    \item Accuracy of learnt x,y and z gradients
    \item Forward looking at the field of INR
\end{enumerate}

Implicit neural representation is a rapidly evolving field of research with many advances occuring within each year, borrowing contemporary deep learning improvements from related fields.
In addition to constant work on improving the high-frequency learning capacity, the inclusion of additional dimensions or
